\section{Introduction}
With the rapid development of smart cities and intelligent transportation systems, vehicular networks have emerged as a crucial component to facilitate the acquisition and analysis of real-time data. Among various applications, video analytics has gained significant attention due to its potential to improve road safety, traffic management, and urban surveillance. However, video analytics tasks are often resource-intensive and require substantial computational power and network bandwidth. As a result, efficiently processing these tasks in vehicular networks presents significant challenges.

Edge computing has become a promising solution to address these challenges by offloading computation-intensive tasks from resource-constrained vehicles to edge servers with stronger processing capabilities. Using edge computing, delay and bandwidth consumption in video analytics tasks can be significantly reduced, improving system performance and task completion rates. In addition, crowdsourcing systems have been widely adopted to utilize diverse vehicular resources by assigning tasks to available vehicles or edge devices.

In this paper, we propose an edge-assisted crowdsourcing system for video analytics tasks in vehicular networks. Our system aims to allocate tasks efficiently to minimize energy consumption while ensuring task completion. The proposed system leverages the computational capabilities of nearby edge servers and vehicle-mounted devices to process video data in a distributed manner.

To achieve efficient task allocation, we model the system by incorporating the characteristics of tasks, workers, and the platform. Each task arriving in the system has specific requirements, such as data size, resolution, and frame rate, which directly impact energy consumption. Workers are modeled as devices with distinct transmission and processing energy profiles, ensuring a detailed representation of real-world energy constraints. The crowdsourcing platform acts as a central controller that assigns tasks to workers while maximizing its own profit by balancing revenue from completed tasks and payment to workers.

The core contribution of this paper lies in the comprehensive modeling framework that integrates multiple factors, including energy consumption, task requirements, and platform revenue. We develop a flexible task allocation strategy that dynamically assigns tasks based on system conditions, ensuring optimal performance. Through extensive simulations, we demonstrate the effectiveness of our proposed system in achieving energy-efficient task allocation while maintaining high task completion rates and maximizing platform profit.

The remainder of this paper is organized as follows: Section II presents the related work, Section III presents the system model, including task modeling, worker modeling, and platform modeling. Section IV introduces the proposed task allocation strategy. Section V evaluates the system's performance through extensive experiments, and Section VI concludes the paper with future research directions.